A Journey from Wild to Textbook Data to Reproducibly Refresh the Wages Data from the National Longitudinal Survey of Youth Database

### 2022-02-15 paper/paper.Rmd
Add subsection calculated experience.

It is conceptually motivated by the variables available in the original data and it also illustrates how one would create a new variable from the information provided by the database, a variable that is more useful than those provided.

@@

## The takeaways

>>>>

Work experience is one of the "key" variables in @SingerJudithD2003Alda as it indicates time variable, instead of using the survey year. Thus, in the spirit of refreshing the data to the newest round of the survey, we also want to include this variable. However,it is not straightforwardly available in the database. Thus, we take the definition of experience mentioned by @SingerJudithD2003Alda to recreate it with the information already provided in the database.

According to @SingerJudithD2003Alda, experience is years after entering the labor force. It identifies the specific day in which individual entering the labor force associated with each wages. Based on that definition, one would assume that experience must be related to the date of individual starting to work. However, after calculating it (`yr_wforce`), we found that the number we obtained was not quite match with the one in @SingerJudithD2003Alda's data. On the other hand, the [topical guide of NLSY79 data](https://www.nlsinfo.org/content/cohorts/nlsy79/topical-guide/employment/work-experience), provided by @nlsy79guide, has the information of variable related to labor force experience. One of the variables mentioned is "number of weeks worked since last interview". We then calculated the experience variables using this information. Since it is only measured since the last interview, to adjust it with the experience since the individual entering the labor force, we calculate the cumulative value. We found a more appropriate results in terms of its closeness with the original data (see Subsection \@ref(takeaways)).

<!--It is conceptually motivated by the variables available in the original data and it also illustrates how one would create a new variable from the information provided by the database, a variable that is more useful than those provided.-->

@@

## The takeaways {#takeaways}

### 2022-02-15 paper/paper.tex
Add subsection calculated experience.

For \texttt{stwork} variable, we only rename the column names from \texttt{new\_data\_qnames}. This variable is then used to calculate the next variable, \texttt{yr\_wforce} for each survey round, which is the year of survey (\texttt{year}) minus the year of individual start working (\texttt{stwork}). Finally, \texttt{exp} variable is derived from the number of weeks worked since the last interview indicated as variable started with \texttt{WKSWK} in \texttt{new\_data\_qnames}. To obtain the work experience since 1979, we calculate the cumulative value. As the measurement unit is in week, we convert this to year. {[}ET: inconsistency in \texttt{st\_work} and \texttt{yr\_wforce}?{]}

@@

It is conceptually motivated by the variables available in the original data and it also illustrates how one would create a new variable from the information provided by the database, a variable that is more useful than those provided.

@@

\hypertarget{the-takeaways}{%
\subsection{The takeaways}\label{the-takeaways}}

>>>>

For \texttt{stwork} variable, we only rename the column names from \texttt{new\_data\_qnames}. This variable is then used to calculate the next variable, \texttt{yr\_wforce} for each survey round, which is the year of survey (\texttt{year}) minus the year of individual start working (\texttt{stwork}). Finally, \texttt{exp} variable is derived from the number of weeks worked since the last interview indicated as variable started with \texttt{WKSWK} in \texttt{new\_data\_qnames}. To obtain the work experience since 1979, we calculate the cumulative value. As the measurement unit is in week, we convert this to year.

@@

Work experience is one of the ``key'' variables in Singer and Willett (2003) as it indicates time variable, instead of using the survey year. Thus, in the spirit of refreshing the data to the newest round of the survey, we also want to include this variable. However,it is not straightforwardly available in the database. Thus, we take the definition of experience mentioned by Singer and Willett (2003) to recreate it with the information already provided in the database.

According to Singer and Willett (2003), experience is years after entering the labor force. It identifies the specific day in which individual entering the labor force associated with each wages. Based on that definition, one would assume that experience must be related to the date of individual starting to work. However, after calculating it (\texttt{yr\_wforce}), we found that the number we obtained was not quite match with the one in Singer and Willett (2003)'s data. On the other hand, the \href{https://www.nlsinfo.org/content/cohorts/nlsy79/topical-guide/employment/work-experience}{topical guide of NLSY79 data}, provided by Bureau of Labor Statistics, U.S. Department of Labor (2021b), has the information of variable related to labor force experience. One of the variables mentioned is ``number of weeks worked since last interview''. We then calculated the experience variables using this information. Since it is only measured since the last interview, to adjust it with the experience since the individual entering the labor force, we calculate the cumulative value. We found a more appropriate results in terms of its closeness with the original data (see Subsection \ref{takeaways}).

@@

\hypertarget{takeaways}{%
\subsection{The takeaways}\label{takeaways}}
